A Simple Minibatch Acceptance Test for MCMC

### 2016-05-20 nips_2016.tex
Changed text to say it is on a log scale

Figure~\ref{fig:gauss_mix_2} (to the left) shows a histogram of the final minibatch sizes used by
the adaptive subsampling method each iteration (capped at 600 for readability).  Usually, the method
can make a decision with 200 samples. Occasionally, however, it must use all 10000 points, resulting
in lots of wasted computation. For this particular experiment, that happened eight times.  In
contrast, our method keeps the minibatch size fixed at 200, but requires that ${\rm std}(\Delta') <
1.2$. Figure~\ref{fig:gauss_mix_2} (in the middle) shows a histogram of the estimated ${\rm
std}(\Delta')$ values each iteration. Only 75 iterations (out of 10000) resulted in ${\rm
std}(\Delta') \ge 1.2$ and for these, we skipped the rest of the iteration (setting
$\theta_{t+1}=\theta_t$). To smooth the ${\rm std}(\Delta')$ values, we use moving average updates.
The third plot in Figure~\ref{fig:gauss_mix_2} investigates the number of times ${\rm std}(\Delta')
\ge 1.2$ based on minibatch sizes. For six sizes (50, 100, 150, 200, 250, and 300), we ran MCMC with
our MH test five times and averaged the number of ``violations.'' As the size increases, the
number of violations decreases, as expected.

>>>>

Figure~\ref{fig:gauss_mix_2} (to the left) shows a histogram of the final minibatch sizes used by
the adaptive subsampling method each iteration (on a log scale for readability).  Usually, the method
can make a decision with 200 samples. Occasionally, however, it must use all 10000 points, resulting
in lots of wasted computation. For this particular experiment, that happened eight times.  In
contrast, our method keeps the minibatch size fixed at 200, but requires that ${\rm std}(\Delta') <
1.2$. Figure~\ref{fig:gauss_mix_2} (in the middle) shows a histogram of the estimated ${\rm
std}(\Delta')$ values each iteration. Only 75 iterations (out of 10000) resulted in ${\rm
std}(\Delta') \ge 1.2$ and for these, we skipped the rest of the iteration (setting
$\theta_{t+1}=\theta_t$). To smooth the ${\rm std}(\Delta')$ values, we use moving average updates.
The third plot in Figure~\ref{fig:gauss_mix_2} investigates the number of times ${\rm std}(\Delta')
\ge 1.2$ based on minibatch sizes. For six sizes (50, 100, 150, 200, 250, and 300), we ran MCMC with
our MH test five times and averaged the number of ``violations.'' As the size increases, the
number of violations decreases, as expected.
